Type-Safe Client-Server Communication in Web Applications

### 2014-02-27 paper.tex
Describe runApp, runClient, Client.

Computations exclusive to the server side live in the \lstinline!Server! monad.
This is basically an IO monad with a few extra operations for session handling;
its main purpose is to prevent the programmer from accidentally attempting to
perform client-exclusive operations, such as popping up a browser dialog box,
on the server.

@@

In order to make an RPC call using this handle, we must supply it with an
argument. This is done using the \lstinline!<.>! operator. Upon closer
inspection, the attentive reader may notice that its type,
\linebreak
\lstinline!Serialize a => Export (a -> b) -> a -> Export b!,
is very similar to the type of the \lstinline!<*>! operator over applicative
functors. This is not a coincidence; the \lstinline!<.>! performs the same role
for the \lstinline!Export! type as \lstinline!<*>! performs for applicative
functors. The reason for using a separate operator for this instead of making
\lstinline!Export! an instance of \lstinline!Applicative! is that since
functions embedded in the \lstinline!Export! type exist only to be called over
a network, such functions must only be applied to arguments which can be
serialized and sent over a network connection.

@@

@@

\section{Discussion and related work}

>>>>

Computations exclusive to the server side live in the \lstinline!Server! monad.
This is basically an IO monad, as can be seen from the regular
\lstinline!putStrLn! \lstinline!IO! computation being lifted into it, with a
few extra operations for session handling; its main purpose is to prevent the
programmer from accidentally attempting to perform client-exclusive operations,
such as popping up a browser dialog box, on the server.

@@

After the \lstinline!export! call, we enter the domain of client-exclusive code
with the application of \lstinline!runClient!. This function, unsurprisingly,
execute computations in the \lstinline!Client! monad which is essentially an
IO monad with cooperative multitasking added on top, to mitigate the fact that
Javascript has no native concurrency support. \lstinline!runClient! does not
return, and is the only function with a return type of \lstinline!App Done!,
which ensures that each \lstinline!App! computation contains exactly one client
computation.

In order to make an RPC call using a handle obtained from \lstinline!export!,
we must supply it with an argument. This is done using the \lstinline!<.>!
operator. Upon closer inspection, the attentive reader may notice that
\linebreak
its type, \lstinline!Serialize a => Export (a -> b) -> a -> Export b!,
is very similar to the type of the \lstinline!<*>! operator over applicative
functors. This is not a coincidence; the \lstinline!<.>! performs the same role
for the \lstinline!Export! type as \lstinline!<*>! performs for applicative
functors. The reason for using a separate operator for this instead of making
\lstinline!Export! an instance of \lstinline!Applicative! is that since
functions embedded in the \lstinline!Export! type exist only to be called over
a network, such functions must only be applied to arguments which can be
serialized and sent over a network connection.

@@

Like \lstinline!runClient!, \lstinline!runApp! does not return, and is intended
to be the only computation executed by \lstinline!main!. In fact, performing
non-deterministic IO and allowing \lstinline!runApp! or its arguments to
capture the result may lead to unpredictable behavior. A remedy for this is
discussed in section \ref{sec:discussion}.

@@

\section{Discussion and related work}\label{sec:discussion}
